Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2021-01-11 mice_synthesizing/paper.Rmd
Typo fix + minor revisions.

A quick approach to lessen the identities in the data is anonimization. However, simply anonimizing the data is not sufficient to fulfill contemporary privacy and confidentiality  requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby reducing the data quality and the scientific return for re-using the same data for further research.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general idea is to use actually observed data to generate multiply imputed synthetic data sets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed data set as a whole with multiple synthetic versions. Alternatively, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available data sets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, such as income values above a certain threshold. [**aren't synthetic values then not also too disclosive?**]

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic datasets allow for correct statistical inferences. The results of the multiple synthetic data sets should be pooled into a single inference, so that the researcher can draw valid conclusions from the pooled results. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality data sets should be kept at a minimum.

To minimize additional effort of creating synthetic datasets on behalf of the researcher, software aimed at multiple imputation of missing data can be employed. Especially if researchers used this software at an earlier stage in the research process, or acquired familiarity with it during earlier projects, the additional burden of creating synthetic datasets is relatively small. The R-package `mice` [@mice] implements multiple imputation of missing data in a straightforward and user-friendly manner. However, the functionality of `mice` is not restricted to the imputation of missing data, but allows to impute any value in the data. Consequently, `mice` can be utilized for the creation of multiply imputed synthetic datasets.

After creating the multiply imputed synthetic data sets, the goal is to obtain valid statistical inferences. In the missing data framework, this is done by performing statistical analyses on all imputed data sets, and pooling the results of the analyses according to Rubin's rules [-@rubin_multiple_1987]. In the synthetic data framework, the same procedure is followed. However, valid statistical results can only be obtained under the premise that the statistical properties of the analyses are correct. Analyzing synthetic data specifically requires appropriate variance estimates. Notably, the variance estimator that is suitable for analyzing multiply imputed datasets for missing data is inappropriate when multiply imputed synthetic datasets are analyzed. As a result, multiple variance estimators have been proposed to make correct inferences from multiply imputed synthetic datasets. These estimators differ according to whether the data consists entirely of synthetic values, or only a subset of the data is synthetic.

>>>>

A quick approach to lessen the identities in the data is anonimization. However, simply anonimizing the data is not sufficient to fulfil contemporary privacy and confidentiality  requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the overarching procedure is to use actually observed data to generate multiply imputed synthetic datasets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed dataset as a whole with multiple synthetic versions. Alternatively, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available datasets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, such as income values above a certain threshold. [**aren't synthetic values then not also too disclosive?**] <!-- I don't think so, because the synthetic values can only be linked to observed values from other sources if they are exactly equal, or if they are the only value within a certain range. The latter might be possible when the number of values is very small, or when the observed value is so extreme that synthetic versions will still be so extreme. However, if the other values are synthetic as well, it is quite likely that synthetic information does not apply to the real-world observation. I think that the highest risk is in the possibility to extract a range of possible values, which might be uniquely match some other record that is available. -->

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic datasets allow for correct statistical inferences, **regardless of the fact that the "true" value is not included in the analyses**. The analyses over multiple synthetic datasets should be pooled into a single inference, so that the researcher can draw valid conclusions from the pooled results. To that respect, the variance should reflect the added variability that is induced by the imputation procedure. 
Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality datasets should be kept at a minimum.

To mitigate additional effort of creating synthetic datasets on behalf of the researcher, software aimed at multiple imputation of missing data can be employed. Especially if researchers used this software at an earlier stage in the research process, or acquired familiarity with it during earlier projects, the additional burden of creating synthetic datasets is relatively small. The R-package `mice` [@mice] implements multiple imputation of missing data in a straightforward and user-friendly manner. However, the functionality of `mice` is not restricted to the imputation of missing data, but allows to impute any value in the data. Consequently, `mice` can be utilized for the creation of multiply imputed synthetic datasets.

After creating the multiply imputed synthetic datasets, the goal is to obtain valid statistical inferences. In the missing data framework, this is done by performing statistical analyses on all imputed datasets, and pooling the results of the analyses according to Rubin's rules [-@rubin_multiple_1987]. In the synthetic data framework, the same procedure is followed. However, valid statistical results can only be obtained under the premise that the statistical properties of the analyses are correct. Analyzing synthetic data specifically requires appropriate variance estimates. Notably, the variance estimator that is suitable for analyzing multiply imputed datasets for missing data is inappropriate when multiply imputed synthetic datasets are analyzed. As a result, multiple variance estimators have been proposed to make correct inferences from multiply imputed synthetic datasets. These estimators differ according to whether the data consists entirely of synthetic values, or only a subset of the data is synthetic.
